---
title: "Join"
author: "M Farhchin"
date: "3/31/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

Combinig data from different tables
murder data 

```{r}
library(tidyverse)
library(dslabs)
data(murders)
head(murders)
```

electoral vote data
```{r}
data(polls_us_election_2016)
head(results_us_election_2016)
```



Note that order of data is not the same in these tables so:
```{r}
identical(results_us_election_2016$state, murders$state)
```

Use left_join (we will remove the others column and rename electoral votes so that the tables fit on the page):

```{r}
tab<-left_join(murders,results_us_election_2016,by='state')%>%
  select(-others)%>%rename(ev='electoral_votes')
head(tab)
```

Plot the results
```{r}
library(ggrepel)
tab %>% ggplot(aes(population/10^6, ev, label = abb)) +
  geom_point() +
  geom_text_repel() + 
  scale_x_continuous(trans = "log2") +
  scale_y_continuous(trans = "log2") +
  geom_smooth(method = "lm", se = FALSE)
```
 Create a subst of from each data set
 
```{r}
# create tab1 object, which is a subset of murders data set
tab1<-slice(murders,1:6)%>%select(state, population)
tab1# create tab2 object which is a subset of eloctoral votes data set
tab2<-slice(results_us_election_2016,c(1:3,5,7:8))%>%
  select(state,electoral_votes)
tab2


```

Use left joint to join these tables but with different orders
```{r}
left_join(tab1,tab2)
```
Use left joint but with different order
```{r}
left_join(tab2,tab1)
```
Use pipe with left join
```{r}
tab1%>%left_join(tab2)
```
Use inner_join to join the table for the rows that are available in both tables
```{r}
inner_join(tab1,tab2)
```
Use full_join
```{r}
full_join(tab1,tab2)
```
Join all elements in tab1 that exist in tab2
```{r}
semi_join(tab1,tab2)
```
Use anti_join to get the elementa that do not exist in tab1 but exist in tab2
```{r}
anti_join(tab1,tab2)
```
Use bind_cols
```{r}
# break tab data intor different columns
tab1<-tab[,1:3] #first 3 columns of tab
tab2<-tab[,4:6] # cols 4 to 6 of tab
tab3<-tab[,7:8] # cols 7 to 9 of tab
# now use bind_cols to combine data
new_tab<-bind_cols(tab1,tab2,tab3)
new_tab
```

Use bind_rows
```{r}
tab1<-tab[1:3,]
tab2<-tab[4:6,]
bind_rows(tab1,tab2)
```
Set operators
```{r}
# apply intersect on vectors
intersect(c(1:5),c(3:10))
```
```{r}
# apply intersect on data frames
r1<-tab[1:5,]
r2<-tab[3:9,]
intersect(r1,r2)
```
```{r}
# apply union on vectors
union(c('a','b','c','c','c','d'),c('a','c','e'))
```

```{r}
# apply union to data frames
union(r1,r2)
```

```{r}
# apply setdiff on vectors
print(setdiff(1:10,5:20))
print(setdiff(5:20,1:10))
```

```{r}
# apply setdiff to data frames
print(setdiff(r1,r2))
print(setdiff(r2,r1))
```
```{r}
setequal(r1,r2)
```

